---
title: "Practical Machine Learning Course Project"
output: html_notebook
---
## Load libraries
```{r}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
```

## Load training data
```{r}
rawData <- read.csv("pml-training.csv")

dim(rawData)

sum(complete.cases(rawData))
```
We can see, that there are nearly no cases without N/A values.
We remove columns containing N/A values:
```{r}
rawData <- rawData[, colSums(is.na(rawData)) == 0]
sum(complete.cases(rawData))
```
We remove "X" column and any other containing timestamp or window in it's name:
```{r}
classe <- rawData$classe
bad_columns <- grepl("^X$|timestamp|window", names(rawData))
dataset <- rawData[, !bad_columns]
dataset <- dataset[, sapply(dataset, is.numeric)]
dataset$classe <- classe

dim(dataset)
```
We see that there are 19622 samples and 53 variables (including the target).

## Split dataset into train and test.

```{r}
set.seed(100)
trainPartition<-createDataPartition(dataset$classe, p=0.70, list=F)
trainData <- dataset[trainPartition,]
testData <- dataset[-trainPartition,]

dim(trainData)
dim(testData)
```

## Model creation
We create a random forest model of 220 trees and train it using 10-fold cross validation.
```{r}
control <- trainControl(method="cv", 10)
model <- train(classe ~., data=trainData, method="rf", ntree=220,trControl=control)
model
```
Estimate the performance of the model on validation set.
```{r}
predictionRf <- predict(model, testData)
confMat <- confusionMatrix(as.factor(testData$classe), predictionRf)

length(predictionRf)
length(testData$classe)
accuracy <- postResample(predictionRf, as.factor(testData$classe))
accuracy
outOfSampleError <- 1 - as.numeric(confMat$overall[1])
outOfSampleError
```
We can see, that the accuracy is 99% and out of sample error 0.7%

## Run model on provided testing data
Prepare the data in the same manner as we have prepared training data before
```{r}
testing <- read.csv("pml-testing.csv")

testing <- testing[, colSums(is.na(testing)) == 0]

bad_columns <- grepl("^X|timestamp|window", names(testing))
testing <- testing[, !bad_columns]
testing <- testing[, sapply(testing, is.numeric)]

dim(testing)
```
We have 20 samples with 44 variables without the ```classe``` variable we want to predict
```{r}
result <- predict(model, testing)
print(result)
```
## Appendix: Figures
```{r}
treeGraph <- rpart(classe ~., data=trainData, method="class")
prp(treeGraph)
```
